---
title: "About priors in INLA"
---

```{css, eval = TRUE, echo = FALSE}
body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
```

Consider the following general code for fitting a model using `inla()` function.

```{r eval = FALSE}
inla(formula = y ~ <covariates> + f("<latent effect>", hyper = hyper),
     family = "<likelihood>", 
     data = <data.frame>,
     control.family = list(hyper = hyper))
```

We assume that `"<likelihood>"` or `"<latent effect>"` have only one hyperparameter called `prec` with default prior `hyper = list(prec = list(prior = "loggamma", param = c(1, 0.00005)), initial = 4, fixed = FALSE)`.


1. If we do

```{r eval = FALSE}
inla(formula = y ~ <covariates> + f("<latent effect>"),
     family = "<likelihood>", 
     data = <data.frame>)
```
 
then the default prior is assigned to the hyperparameter and estimation is performed.
 
2. If we set the `hyper` parameter to

    - `hyper = list(prec = list(initial = 0, fixed = TRUE))`, then the hyperparameter `prec` is fixed to its initial value `initial`, estimation is not performed, and the prior distribution is ignored. **Neither the default prior nor its parameters `param` are modified**.

    - `hyper = list(prec = list(initial = 0, fixed = FALSE))`, then the hyperparameter `prec` is initialized to `initial` and estimation is performed. **Neither the default prior nor its parameters `param` are modified**.

    - `hyper = list(prec = list(prior = "pc.prec", param = c(5, 0.01)))`, then the prior is totally changed.

3. Run `inla.doc("<llkelihood>")` or `inla.doc("<latent effect>")` to get more information.


