---
title: "Highways14"
author: "Several authors"
date: "Created: 12-04-2024. Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show
    # df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
always_allow_html: true
---

```{r, eval = FALSE, echo = FALSE}
################################################################################
################################################################################
################################################################################
################################################################################
######### DO NOT FORGET TO CHANGE THE TITLE EVERY TIME YOU FIT A MODEL #########
################################################################################
################################################################################
################################################################################
################################################################################
```

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
```


Let us set some global options for all code chunks in this document.


```{r}
knitr::opts_chunk$set(
  message = FALSE,    # Disable messages printed by R code chunks
  warning = FALSE,    # Disable warnings printed by R code chunks
  echo = TRUE,        # Show R code within code chunks in output
  include = TRUE,     # Include both R code and its results in output
  eval = TRUE,       # Evaluate R code chunks
  cache = FALSE,       # Enable caching of R code chunks for faster rendering
  fig.align = "center",
  out.width = "100%",
  retina = 2,
  error = TRUE,
  collapse = TRUE
)
rm(list = ls())
set.seed(1982)
```

# PREPROCESSING

Let us now load some required libraries.

```{r}
# Load required libraries

# inla.upgrade(testing = TRUE)
# remotes::install_github("inlabru-org/inlabru", ref = "devel")
# remotes::install_github("davidbolin/rspde", ref = "devel")
# remotes::install_github("davidbolin/metricgraph", ref = "devel")
# remotes::install_github("davidbolin/ngme2", ref = "devel")

library(INLA)
#inla.setOption(num.threads = 7)
library(inlabru)
library(rSPDE)
library(MetricGraph)
library(ngme2)

library(plotly)
library(dplyr)

library(sf)

library(here)
```

Function `standarize()` below is later used to standardize the covariate `SpeedLimit`.

```{r}
standardize <- function(x) {return((x - mean(x)) / sd(x))}
```



## PREPARING AND ADDING THE DATA TO THE GRAPH.


We load the graph object `sf_graph` (which only contains weights) and the data (already graph-processed). 

```{r}
timeallprocedure <- Sys.time()
load(here("Graph_objects/graph_construction_19MAY24_FRC0134.RData"))
load(here("Data_files/data_day7142128_hour13_with_no_consecutive_zeros_19MAY24_FRC0134_graph_processed.RData"))
data_on_graph = data_on_graph %>% 
  dplyr::select(-datetime)
```


The following commands remove zero speed observations that are 1m away from the graph, and after that, they remove any speed observations that are 3m away from the graph.

```{r}
to_remove = data_on_graph %>%
  filter(speed == 0, .distance_to_graph > 0.001) 

data_on_graph = setdiff(data_on_graph, to_remove) %>% 
  filter(.distance_to_graph <= 0.003)
```


We add data to the graph.

```{r}
sf_graph$add_observations(data = data_on_graph, 
                          group = "day", 
                          normalized = TRUE, 
                          clear_obs = TRUE)
```

We get the values of the weights at data locations. This essentially gives us covariates from the weights.

```{r}
sf_graph$edgeweight_to_data(data_loc = TRUE)
```


When running `sf_graph$edgeweight_to_data(data_loc = TRUE)`, some `NA` values are created (because the data is grouped). We remove them below. We also standardize the `SpeedLimit` covariate. 

```{r}
data <- sf_graph$get_data() %>% 
  drop_na(-StreetName) %>% # this drops all rows with at least one NA value but without taking into account StreetName
  mutate(across(c("SpeedLimit"), ~standardize(.))) %>%
  dplyr::select(speed, SpeedLimit)
```

## NOW WE ADD THE FINAL VERSION OF THE DATA TO THE GRAPH

We add the data again but now with the new standardized `SpeedLimit` covariate.

```{r}
sf_graph$add_observations(data = data, 
                          group = "day", 
                          normalized = TRUE, 
                          clear_obs = TRUE)
```

## GET THE OBSERVATION LOCATIONS (FROM THE (FINAL VERSION OF THE) DATA EXTRACTED FROM THE GRAPH)

```{r}
data_final <- sf_graph$get_data()
obs.loc <- data_final |> as.data.frame() |> dplyr::select(.edge_number, .distance_on_edge)
```


## BUILD THE MESH

```{r}
h = 0.05
sf_graph$build_mesh(h = h)
```

## GET THE VALUES OF THE COVARIATE ON THE MESH LOCATIONS

We get the value of the weights at mesh locations. This will allow us to built matrices `B.sigma` and `B.range` below. Again, `sf_graph$edgeweight_to_data(mesh = TRUE, add = FALSE, return = TRUE)` creates repeated information (because the data is grouped). We fix that by filtering one group. We also standardize the `SpeedLimit` covariate. 

```{r}
mesh <- sf_graph$edgeweight_to_data(mesh = TRUE, 
                                   add = FALSE, 
                                   return = TRUE) %>% 
  filter(.group == 1) %>%
  mutate(across(c("SpeedLimit"), ~standardize(.))) %>%
  dplyr:::select.data.frame(SpeedLimit)
```



```{r}
res <- lm(speed ~ SpeedLimit,
         data = data_final)

summary(res)

yhat <- res$fitted.values
plot(yhat)
```

```{r}
plot(res)
```

```{r}
plot(data_final$SpeedLimit, data_final$speed)
abline(res, col = "red", lwd = 2)

```



```{r}
residuals <- res$residuals
residuals2 <- data_final$speed - yhat
```

```{r}
data_final <- data_final %>%
  mutate(residuals = residuals2,
         yhat = yhat)
```


```{r}
ggplot() +
  geom_point(data = data_final, aes(x = .coord_x, y = .coord_y, color = speed)) +
  scale_color_viridis_c()
ggplot() +
  geom_point(data = data_final, aes(x = .coord_x, y = .coord_y, color = SpeedLimit)) +
  scale_color_viridis_c()
ggplot() +
  geom_point(data = data_final, aes(x = .coord_x, y = .coord_y, color = yhat)) +
  scale_color_viridis_c()
ggplot() +
  geom_point(data = data_final, aes(x = .coord_x, y = .coord_y, color = residuals)) +
  scale_color_viridis_c()
```



```{r}
save.image(here(paste0("Models_output/", rmarkdown::metadata$title, ".RData")))
```



