---
title: "Model12"
author: "Several authors"
date: "Created: 12-04-2024. Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show
    # df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
always_allow_html: true
---

```{r, eval = FALSE, echo = FALSE}
################################################################################
################################################################################
################################################################################
################################################################################
######### DO NOT FORGET TO CHANGE THE TITLE EVERY TIME YOU FIT A MODEL #########
################################################################################
################################################################################
################################################################################
################################################################################
```

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
```


Let us set some global options for all code chunks in this document.


```{r}
knitr::opts_chunk$set(
  message = FALSE,    # Disable messages printed by R code chunks
  warning = FALSE,    # Disable warnings printed by R code chunks
  echo = TRUE,        # Show R code within code chunks in output
  include = TRUE,     # Include both R code and its results in output
  eval = TRUE,       # Evaluate R code chunks
  cache = FALSE,       # Enable caching of R code chunks for faster rendering
  fig.align = "center",
  out.width = "100%",
  retina = 2,
  error = TRUE,
  collapse = TRUE
)
rm(list = ls())
set.seed(1982)
```

# Preprocessing

Let us now load some required libraries.

```{r}
# Load required libraries

# inla.upgrade(testing = TRUE)
# remotes::install_github("inlabru-org/inlabru", ref = "devel")
# remotes::install_github("davidbolin/rspde", ref = "devel")
# remotes::install_github("davidbolin/metricgraph", ref = "devel")

library(INLA)
inla.setOption(num.threads = 7)
library(inlabru)
library(rSPDE)
library(MetricGraph)

library(plotly)
library(dplyr)

library(sf)

library(here)
```

Function `standarize()` below is later used to standardize the covariate `SpeedLimit`.

```{r}
standardize <- function(x) {return((x - mean(x)) / sd(x))}
```



-------------------

**To keep track of the changes, we provide summaries of every new created object. Those summaries can be accessed by pressing the Show buttons below**

------------------


We load the graph object `sf_graph` (which only contains weights) and the data (already graph-processed). 

```{r}
load(here("Graph_objects/graph_construction_25_04_2024partialtomtomwhichlonglatsf.RData"))
load(here("Data_files/data_day7142128_hour13_with_no_consecutive_zeros_partialtomtom_graph_25_04_2024_processed.RData"))
data_on_graph = data_on_graph %>% 
  dplyr::select(-datetime)
```

```{r}
sf_graph$get_edge_lengths() %>% head() %>% capture.output() %>% grep("^Units:", ., value = TRUE)
```


```{r, class.source = "fold-hide"}
summary(sf_graph)
summary(data_on_graph)
```

The following commands remove zero speed observations that are 1m away from the graph, and after that, they remove any speed observations that are 3m away from the graph.

```{r}
to_remove = data_on_graph %>%
  filter(speed == 0, .distance_to_graph > 0.001) 

data_on_graph = setdiff(data_on_graph, to_remove) %>% 
  filter(.distance_to_graph <= 0.003)
```


```{r, class.source = "fold-hide"}
summary(to_remove)
summary(data_on_graph)
```

We add data to the graph.

```{r}
sf_graph$add_observations(data = data_on_graph, 
                          group = "day", 
                          normalized = TRUE, 
                          clear_obs = TRUE)
```


```{r, class.source = "fold-hide"}
sf_graph$get_data()
summary(sf_graph)
```

We get the values of the weights at data locations. This essentially gives us covariates from the weights.

```{r}
sf_graph$edgeweight_to_data(data_loc = TRUE)
```


```{r, class.source = "fold-hide"}
sf_graph$get_data()
summary(sf_graph)
```

When running `sf_graph$edgeweight_to_data(data_loc = TRUE)`, some `NA` values are created (because the data is grouped). We remove them below. We also standardize the `SpeedLimit` covariate. 

```{r}
data = sf_graph$get_data() %>% 
  drop_na(-StreetName) %>% # this drops all rows with at least one NA value but without taking into account StreetName
  mutate(across(c("SpeedLimit"), ~standardize(.))) %>%
  dplyr::select(speed, SpeedLimit)
```


--- 

The code of chunk below was executed only one time.

---

```
{r, eval = FALSE}
aux = data |>
  rename(distance_on_edge = .distance_on_edge, edge_number = .edge_number) |>
  as.data.frame() |>
  dplyr::select(edge_number, distance_on_edge, .group)

distmatrixlist = list()

for (i in 1:4) {
  distmatrixlist[[i]] = sf_graph$compute_geodist_PtE(PtE = aux %>% 
                                                       filter(.group == as.character(i)) %>% 
                                                       dplyr::select(-.group),
                                                     normalized = TRUE,
                                                     include_vertices = FALSE)
}


save(distmatrixlist, file = here("Models_output/distmatrixfixed.RData"))
```

--- 

The code of chunk above was executed only one time.

---

```{r, class.source = "fold-hide"}
summary(data)
```

We add the data again but now with the new standardized `SpeedLimit` covariate.

```{r}
sf_graph$add_observations(data = data, 
                          group = "day", 
                          normalized = TRUE, 
                          clear_obs = TRUE)
```


```{r, class.source = "fold-hide"}
sf_graph$get_data()
summary(sf_graph)
```

We build a mesh.

```{r}
h = 0.05
sf_graph$build_mesh(h = h)
```


```{r, class.source = "fold-hide"}
summary(sf_graph)
```

We get the value of the weights at mesh locations. This will allow us to built matrices `B.sigma` and `B.range` below. Again, `sf_graph$edgeweight_to_data(mesh = TRUE, add = FALSE, return = TRUE)` creates repeated information (because the data is grouped). We fix that by filtering one group. We also standardize the `SpeedLimit` covariate. 

```{r}
mesh = sf_graph$edgeweight_to_data(mesh = TRUE, 
                                   add = FALSE, 
                                   return = TRUE) %>% 
  filter(.group == 1) %>%
  mutate(across(c("SpeedLimit"), ~standardize(.))) %>%
  dplyr:::select.data.frame(SpeedLimit)
```


```{r, class.source = "fold-hide"}
summary(mesh)
```


:::: {style="display: flex; to display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}

::: {}

## Stationary model


- Observe that we are considering replicates.

```{r}
stat.time.ini <- Sys.time()
################################################################################
################################# STATIONARY MODEL #############################
################################################################################

rspde_model_stat <- rspde.metric_graph(sf_graph,
                                         parameterization = "matern",
                                         nu.upper.bound = 1.5)
```


```{r, class.source = "fold-hide"}
str(rspde_model_stat)
```


```{r}
data_rspde_bru_stat <- graph_data_rspde(rspde_model_stat,
                                        repl = ".all",
                                        loc_name = "loc")

```


```{r, class.source = "fold-hide"}
str(data_rspde_bru_stat)
```




```{r}
cmp_stat = speed ~ -1 +
  Intercept(1) +
  field(loc, model = rspde_model_stat,
        replicate = data_rspde_bru_stat[["repl"]])

rspde_fit_stat <-
  bru(cmp_stat,
      data = data_rspde_bru_stat[["data"]],
      family = "gaussian",
      options = list(verbose = FALSE)
  )
```


```{r, class.source = "fold-hide"}
str(rspde_fit_stat)
```


```{r, collapse = FALSE}
stat.time.fin <- Sys.time()
print(stat.time.fin - stat.time.ini)

summary(rspde_fit_stat)
fit.rspde = rspde.result(rspde_fit_stat, "field", rspde_model_stat)
summary(fit.rspde)
```

:::

::: {}

## Nonstationary model

- Observe that we are using the computed parameters from the stationary model as initial values for the nonstationary models.

```{r}
nonstat.time.ini <- Sys.time()
################################################################################
############################# NON STATIONARY MODEL #############################
################################################################################

B.sigma = cbind(0, 1, 0, mesh$SpeedLimit, 0)
B.range = cbind(0, 0, 1, 0, mesh$SpeedLimit)
init.vec.theta = c(fit.rspde$summary.log.std.dev$mode, 
                   fit.rspde$summary.log.range$mode, 
                   rep(0, (ncol(B.sigma)-3)))

rspde_model_nonstat <- rspde.metric_graph(sf_graph,
                                          start.theta = init.vec.theta,
                                          theta.prior.mean = init.vec.theta,
                                          B.sigma = B.sigma,
                                          B.range = B.range,
                                          parameterization = "matern",
                                          nu.upper.bound = 1.5)
```


```{r, class.source = "fold-hide"}
str(rspde_model_nonstat)
```


```{r}
data_rspde_bru_nonstat <- graph_data_rspde(rspde_model_nonstat,
                                           repl = ".all",
                                           loc_name = "loc")
```


```{r, class.source = "fold-hide"}
str(data_rspde_bru_nonstat)
```


```{r}
cmp_nonstat = speed ~ -1 +
  Intercept(1) +
  field(loc, model = rspde_model_nonstat,
        replicate = data_rspde_bru_nonstat[["repl"]])

rspde_fit_nonstat <-
  bru(cmp_nonstat,
      data = data_rspde_bru_nonstat[["data"]],
      family = "gaussian",
      options = list(verbose = FALSE)
  )
```


```{r, class.source = "fold-hide"}
str(rspde_fit_nonstat)
```


```{r collapse = FALSE}
nonstat.time.fin <- Sys.time()
print(nonstat.time.fin - nonstat.time.ini)

summary(rspde_fit_nonstat)
summary(rspde.result(rspde_fit_nonstat, "field", rspde_model_nonstat))

```

:::

::::



## Crossvalidation 1

```{r}
#load(here("Models_output/distmatrixfixed.RData"))

points = data %>%
  as.data.frame() %>%
  st_as_sf(coords = c(".coord_x", ".coord_y"), crs = 4326) %>%
  mutate(., index = 1:nrow(.)) %>% 
  st_drop_geometry() %>%
  dplyr:::select(speed, .group, index) %>%
  mutate(.group = as.numeric(.group)) %>%
  group_by(.group) %>%
  mutate(indexingroup = seq_len(n())) %>%
  ungroup()

distance = seq(from = 0, to = 400, by = 20)/1000

#save(GROUPS, file = here("Models_output/GROUPS_for_window_case.RData"))
```

```{r, fig.height = 8}
load(here("Models_output/GROUPS_for_window_case.RData"))
mse.stat <- mse.nonstat <- ls.stat <- ls.nonstat <- rep(0,length(distance))
# cross-validation for-loop
for (j in 1:length(distance)) {
  print(j)
  # cross-validation of the stationary model
  cv.stat <- inla.group.cv(rspde_fit_stat, groups = GROUPS[[j]])
  # cross-validation of the nonstationary model
  cv.nonstat <- inla.group.cv(rspde_fit_nonstat, groups = GROUPS[[j]])
  # obtain MSE and LS
  mse.stat[j] <- mean((cv.stat$mean - points$speed)^2)
  mse.nonstat[j] <- mean((cv.nonstat$mean - points$speed)^2)
  ls.stat[j] <- mean(log(cv.stat$cv))
  ls.nonstat[j] <- mean(log(cv.nonstat$cv))
}

## plot results
par(mfrow = c(2,2), family = "Palatino")

# Plot MSE
plot(distance, mse.stat, main = "MSE", ylim = c(min(mse.nonstat, mse.stat), max(mse.nonstat, mse.stat)),
     type = "l", ylab = "MSE", xlab = "distance in m", col = "black")
lines(distance, mse.nonstat, col = "blue")
legend("bottomright", legend = c("Stationary", "Non-stationary"), col = c("black", "blue"), lty = 1)

# Plot log-score
plot(distance, -ls.stat, main = "log-score", ylim = c(min(-ls.nonstat, -ls.stat), max(-ls.nonstat, -ls.stat)),
     type = "l", ylab = "log-score", xlab = "distance in m", col = "black")
lines(distance, -ls.nonstat, col = "blue")
legend("bottomright", legend = c("Stationary", "Non-stationary"), col = c("black", "blue"), lty = 1)

```



```{r}
save.image(here(paste0("Models_output/", rmarkdown::metadata$title, ".RData")))
```

