---
title: "Model7"
author: "Several authors"
date: "Created: 12-04-2024. Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
always_allow_html: true
---

```{r, eval = FALSE, echo = FALSE}
################################################################################
################################################################################
################################################################################
################################################################################
######### DO NOT FORGET TO CHANGE THE TITLE EVERY TIME YOU FIT A MODEL #########
################################################################################
################################################################################
################################################################################
################################################################################
```

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
```


Let us set some global options for all code chunks in this document.


```{r}
knitr::opts_chunk$set(
  message = FALSE,    # Disable messages printed by R code chunks
  warning = FALSE,    # Disable warnings printed by R code chunks
  echo = TRUE,        # Show R code within code chunks in output
  include = TRUE,     # Include both R code and its results in output
  eval = TRUE,       # Evaluate R code chunks
  cache = FALSE,       # Enable caching of R code chunks for faster rendering
  fig.align = "center",
  out.width = "100%",
  retina = 2,
  error = TRUE
)
rm(list = ls())
set.seed(1982)
```

# Preprocessing

Let us now load some required libraries.

```{r}
# Install or upgrade required libraries

# inla.upgrade(testing = TRUE)
# remotes::install_github("inlabru-org/inlabru", ref = "devel")
# remotes::install_github("davidbolin/rspde", ref = "devel")
# remotes::install_github("davidbolin/metricgraph", ref = "devel")

# Load required libraries

library(INLA)
library(inlabru)
library(rSPDE)
library(MetricGraph)

library(plotly)
library(dplyr)

library(here)
```

We now define some `R` functions with self-explanatory names.

```{r}
# Functions to manipulate the data
standardize <- function(vector) {return((vector - mean(vector)) / sd(vector))}
minmax_scaling <- function(x) {return((x - min(x)) / (max(x) - min(x)))}
convert_to_binary <- function(input_vector) {return(ifelse(input_vector != 0, 1, 0))}
```

### Prepare the data

The following chunk uses the function `creates_covariates_on_mesh()` defined on `Preprocessing/28Window.case.file4.addcovariatesonmesh.R` to build a mesh (on the same graph as the one used in this file) and compute the value of all available covariates on the mesh nodes.

Go to [**Preprocessing 8.0.1 Raw code**](dp.html#801_Raw_code) to see the body of the `creates_covariates_on_mesh()` function.


```{r}
h = 0.05
# source(here("Preprocessing/28Window.case.file4.addcovariatesonmesh.R"))
# creates_covariates_on_mesh(h)
# Sys.sleep(5)

################################################################################
################################# PREPARE THE DATA #############################
################################################################################

# loading the data
load(here("Data_files/data_on_graph_with_covariates_no_consecutive_zeros_partialtomtom.RData"))
load(here("Graph_objects/graph_construction_28_03_2024partialtomtomwhichlonglatsf.RData"))
load(here("Data_files/data_on_mesh_with_covariates_partialtomtom.RData"))


data = data_on_graph_with_covariates %>%
  mutate(across(starts_with(c("class_", "upto")), list(ind = convert_to_binary))) %>%
  mutate(across(c("bus", "signal", "stop", "crossing"), ~round(., 5))) %>%
  mutate(across(c("SpeedLimit", "density_per_hour"), ~standardize(.))) %>% dplyr::select(-datetime)


mesh = data_on_mesh_with_covariates %>%
  mutate(across(starts_with(c("class_", "upto")), list(ind = convert_to_binary))) %>% # this creates new columns
  mutate(across(c("bus", "signal", "stop", "crossing"), ~round(., 5))) %>%
  mutate(across(c("SpeedLimit", "density_per_hour"), ~standardize(.)))
```

### Add data and build mesh

Let us add the speed observations to the graph and build the mesh. Observe that `mesh` already has the value of the available covariates on the mesh nodes.

```{r}
sf_graph$add_observations(data = data, group = "day", normalized = TRUE, clear_obs = TRUE)
sf_graph$build_mesh(h = h)
Sys.sleep(5)
```



# Modeling

### Stationary model

- Observe that we are considering replicates.

```{r}
stat.time.ini <- Sys.time()
################################################################################
################################# STATIONARY MODEL #############################
################################################################################

rspde_model_stat <- rspde.metric_graph(sf_graph,
                                         parameterization = "matern",
                                         nu = 0.5)

data_rspde_bru_stat <- graph_data_rspde(rspde_model_stat,
                                        repl = ".all",
                                        name = "field")

st.dat.rep.stat <- inla.stack(
    data = data_rspde_bru_stat[["data"]]["speed"], 
    A = list(data_rspde_bru_stat[["basis"]],1), 
    tag = "est",
    effects =
      list(c(data_rspde_bru_stat[["index"]], list(Intercept = 1)), 
           list(SpeedLimit = data_rspde_bru_stat[["data"]]["SpeedLimit"])
      )
    )


f_stat = speed ~ -1 + Intercept + SpeedLimit + f(field, model = rspde_model_stat, replicate = field.repl)

rspde_fit_stat <-
  inla(f_stat,
      data = inla.stack.data(st.dat.rep.stat),
      family = "gaussian",
      control.predictor = list(A = inla.stack.A(st.dat.rep.stat))
  )

stat.time.fin <- Sys.time()
print(stat.time.fin - stat.time.ini)

summary(rspde_fit_stat)
fit.rspde = rspde.result(rspde_fit_stat, "field", rspde_model_stat)
summary(fit.rspde)
```


### Nonstationary model

- Observe that we are using the computed parameters from the stationary model as initial values for the nonstationary models.

```{r}
nonstat.time.ini <- Sys.time()
################################################################################
############################# NON STATIONARY MODEL #############################
################################################################################

B.sigma = cbind(0, 1, 0, mesh$SpeedLimit, 0)
B.range = cbind(0, 0, 1, 0, mesh$SpeedLimit)

rspde_model_nonstat <- rspde.metric_graph(sf_graph,
                                            start.theta = c(fit.rspde$summary.log.std.dev$mode,
                                                            fit.rspde$summary.log.range$mode,
                                                            rep(0, 2)),
                                            B.sigma = B.sigma,
                                            B.range = B.range,
                                            parameterization = "matern",
                                            nu = 0.5)

data_rspde_bru_nonstat <- graph_data_rspde(rspde_model_nonstat,
                                        repl = ".all",
                                        name = "field")

st.dat.rep.nonstat <- inla.stack(
    data = data_rspde_bru_nonstat[["data"]]["speed"], 
    A = list(data_rspde_bru_nonstat[["basis"]],1), 
    tag = "est",
    effects =
      list(c(data_rspde_bru_nonstat[["index"]], list(Intercept = 1)), 
           list(SpeedLimit = data_rspde_bru_nonstat[["data"]]["SpeedLimit"])
      )
    )


f_nonstat = speed ~ -1 + Intercept + SpeedLimit + f(field, model = rspde_model_nonstat, replicate = field.repl)

rspde_fit_nonstat <-
  inla(f_nonstat,
      data = inla.stack.data(st.dat.rep.nonstat),
      family = "gaussian",
      control.predictor = list(A = inla.stack.A(st.dat.rep.nonstat))
  )

nonstat.time.fin <- Sys.time()
print(nonstat.time.fin - nonstat.time.ini)

summary(rspde_fit_nonstat)
summary(rspde.result(rspde_fit_nonstat, "field", rspde_model_nonstat))

```



```{r}
save.image(here(paste0("Models_output/", rmarkdown::metadata$title, ".RData")))
```

