---
title: "Model10"
author: "Several authors"
date: "Created: 12-04-2024. Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show
    # df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
always_allow_html: true
---

```{r, eval = FALSE, echo = FALSE}
################################################################################
################################################################################
################################################################################
################################################################################
######### DO NOT FORGET TO CHANGE THE TITLE EVERY TIME YOU FIT A MODEL #########
################################################################################
################################################################################
################################################################################
################################################################################
```

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
  font-family: 'Fira Code', 'Lucida Console', Monaco, monospace;
}
pre {
  font-size: 14px
}
```


```{r,echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,    # Disable messages printed by R code chunks
  warning = FALSE,    # Disable warnings printed by R code chunks
  echo = TRUE,        # Show R code within code chunks in output
  include = TRUE,     # Include both R code and its results in output
  eval = FALSE,       # Evaluate R code chunks
  cache = TRUE,       # Enable caching of R code chunks for faster rendering
  fig.align = "center",
  out.width = "100%",
  retina = 2,
  error = TRUE,
  collapse = TRUE
)
rm(list = ls())
set.seed(1982)
```

```{r, echo = FALSE, eval = TRUE}
# Load required libraries

# inla.upgrade(testing = TRUE)
# remotes::install_github("inlabru-org/inlabru", ref = "devel")
# remotes::install_github("davidbolin/rspde", ref = "devel")
# remotes::install_github("davidbolin/metricgraph", ref = "devel")

library(INLA)
library(inlabru)
library(rSPDE)
library(MetricGraph)

library(plotly)
library(dplyr)

library(sf)

library(here)
```


# Structure

The basic structure of a `metric_graph` object is defined as

```{r}
metric_graph <-  R6Class(classname = "metric_graph", 
                         public = list(),
                         private = list(line_to_vertex = function(args),
                                        PtE_to_mesh = function(args),
                                        plot_2d = function(args),
                                        plot_3d = function(args),
                                        coordinates_multiple_snaps = function(args),
                                        merge_close_vertices = function(args),
                                        remove_circles = function(args),
                                        merge.all.deg2 = function(),
                                        remove.first.deg2 = function(args),
                                        compute_lengths = function(args),
                                        A = function(args),
                                        get_edge_weights_internal = function(args),
                                        clear_initial_info = function(),
                                        split_edge = function(args),
                                        compute_laplacian_PtE = function(args),
                                        find_edge_edge_points = function(arga),
                                        add_vertices = function(args),
                                        compute_degrees = function(),
                                        create_update_vertices = function(),
                                        mesh_merge_deg2 = function(),
                                        mesh_merge_outs = function(),
                                        move_V_first = function(),
                                        find_mesh_bc = function(),
                                        set_petrov_matrices = function(),
                                        set_first_weights = function(args)
                                        )
                         )
```

It only considers `classname`, `public`, and `private` arguments and does not consider any of the other available arguments such as `active` or `inherit`. See below.

```{r}
R6Class(
  classname = NULL,
  public = list(),
  private = NULL,
  active = NULL,
  inherit = NULL,
  lock_objects = TRUE,
  class = TRUE,
  portable = TRUE,
  lock_class = FALSE,
  cloneable = TRUE,
  parent_env = parent.frame(),
  lock
)
```

# Some nuances

- It is not a coincidence that the `classname` parameter coincides with the variable name of the output of `R6Class()`.

- The `$new()` method when we do `metric_graph$new()` is not defined by the authors but a default one. The `$new()` method creates the object and calls the `initialize()` method, if it exists.

# Unknown (possibly private) functions

- `process_factor_unit()`

- `private$set_first_weights()`

- `private$line_to_vertex()`

- `private$find_edge_edge_points()`


# `initialize()` inputs

```{r}
initialize = function(edges = NULL,
                      V = NULL,
                      E = NULL,
                      vertex_unit = NULL,
                      length_unit = vertex_unit,
                      edge_weights = 1,
                      kirchhoff_weights = NULL,
                      longlat = FALSE,
                      crs = NULL,
                      proj4string = NULL,
                      which_longlat = "sp",
                      project = FALSE,
                      project_data = FALSE,
                      which_projection = "Winkel tripel",
                      tolerance = list(vertex_vertex = 1e-3,
                                       vertex_edge = 1e-3,
                                       edge_edge = 0),
                      check_connected = TRUE,
                      remove_deg2 = FALSE,
                      merge_close_vertices = TRUE,
                      factor_merge_close_vertices = 1,
                      remove_circles = TRUE,
                      verbose = 1,
                      lines = deprecated()) {
  # function body
}
```


# `lines` argument

---

`lifecycle::is_present()` and `lifecycle::deprecate_warn()`

---

```{r}
if (lifecycle::is_present(lines)) {
         if (is.null(edges)) {
           lifecycle::deprecate_warn("1.2.0", "metric_graph$new(lines)", "metric_graph$new(edges)",
             details = c("`lines` was provided but not `edges`. Setting `edges <- lines`.")
           )
           edges <- lines
         } else {
           lifecycle::deprecate_warn("1.2.0", "metric_graph$new(lines)", "metric_graph$new(edges)",
             details = c("Both `edges` and `lines` were provided. Only `edges` will be considered.")
           )
         }
         lines <- NULL
       }
```

- Function `lifecycle::is_present()` checks if `lines` was provided by user.

- Function `lifecycle::deprecate_warn()` gives a nice formatted warning.

```{r, class.source = "fold-hide"}
foobar_adder <- function(foo, bar, baz = deprecated()) {
  # Check if user has supplied `baz` instead of `bar`
  if (lifecycle::is_present(baz)) {

    # Signal the deprecation to the user
    deprecate_warn("1.0.0", "foo::bar_adder(baz = )", "foo::bar_adder(bar = )")

    # Deal with the deprecated argument for compatibility
    bar <- baz
  }

  foo + bar
}

foobar_adder(1, 2)
#> [1] 3
foobar_adder(1, baz = 2)
#> Warning: The `baz` argument of `bar_adder()` is deprecated as of foo 1.0.0.
#> ℹ Please use the `bar` argument instead.
#> [1] 3
```


```{r, eval = TRUE, message = TRUE, warning  = TRUE, class.source = "fold-hide"}
data_frame <- function(...) {
  lifecycle::deprecate_warn("1.1.0", "data_frame()", "tibble()")
  tibble::tibble(...)
}
df1 <- data_frame(x = 1, y = 2)
```


# `kirchhoff_weights` argument

---

`stop()`

---

::: {.alert .alert-info}
Here we start to define `private` fields, which allows to create fields and methods that are only available from within the class, not outside of it. Visit [here](https://adv-r.hadley.nz/r6.html) and [here](https://r6.r-lib.org/articles/Introduction.html) for more information.
:::

```{r}
if(!is.null(kirchhoff_weights)){
        if(length(kirchhoff_weights)>1){
          warning("Only the first entry of 'kirchhoff_weights' was used.")
          kirchhoff_weights <- kirchhoff_weights[[1]]
        }
        if(!is.numeric(kirchhoff_weights)){
          if(!is.character(kirchhoff_weights)){
            stop("'kirchhoff_weights' must be either a number of a string.")
          }
          if(!(kirchhoff_weights%in%colnames(edge_weights))){
            stop(paste(kirchhoff_weights, "is not a column of 'edge_weights'!"))
          }
        } else{
          if(!is.data.frame(edge_weights)){
            if(kirchhoff_weights != 1){
              stop("Since 'edge_weights' is not a data.frame, 'kirchhoff_weights' must be either NULL or 1.")
            }
          } else{
            if(kirchhoff_weights %%1 != 0){
              stop("'kirchhoff_weights' must be an integer.")
            }
            if((kirchhoff_weights < 1) || (kirchhoff_weights > ncol(edge_weights))){
              stop("'kirchhoff_weights' must be a positive integer number smaller or equal to the number of columns of 'edge_weights'.")
            }
          }
        }
        private$kirchhoff_weights <- kirchhoff_weights
       }
```

- Function `stop()` stops the execution of the entire function at the point where it is located.

```{r, eval = TRUE}
sqrt_non_negative <- function(x) {
  print(10)
  if (x < 0) {
    stop("Input must be non-negative.")
  }
  print(25)
  return(sqrt(x))
}

sqrt_non_negative(4)   
sqrt_non_negative(-1)  

```

# `tolerance` argument

```{r}
if (is.null(tolerance$vertex_edge) && !is.null(tolerance$vertex_line)) {
           lifecycle::deprecate_warn("1.2.0", "metric_graph$new(tolerance = 'must contain either vertex_vertex, vertex_edge or edge_edge')",
             details = c("`tolerance$vertex_line` was provided but not `tolerance$vertex_edge`. Setting `tolerance$vertex_edge <- tolerance$vertex_line`.")
           )
           tolerance$vertex_edge <- tolerance$vertex_line
           tolerance$vertex_line <- NULL
         } else if(!is.null(tolerance$vertex_edge) && !is.null(tolerance$vertex_line)) {
           lifecycle::deprecate_warn("1.2.0","metric_graph$new(tolerance = 'must contain either vertex_vertex, vertex_edge or edge_edge')",
             details = c("Both `tolerance$vertex_edge` and `tolerance$vertex_line` were provided. Only `tolerance$vertex_edge` will be considered.")
           )
            tolerance$vertex_line <- NULL
         }

        if (is.null(tolerance$edge_edge) && !is.null(tolerance$line_line)) {
           lifecycle::deprecate_warn("1.2.0", "metric_graph$new(tolerance = 'must contain either vertex_vertex, vertex_edge or edge_edge')",
             details = c("`tolerance$line_line` was provided but not `tolerance$edge_edge`. Setting `tolerance$edge_edge <- tolerance$line_line`.")
           )
           tolerance$edge_edge <- tolerance$line_line
           tolerance$line_line <- NULL
         } else if(!is.null(tolerance$edge_edge) && !is.null(tolerance$line_line)) {
           lifecycle::deprecate_warn("1.2.0","metric_graph$new(tolerance = 'must contain either vertex_vertex, vertex_edge or edge_edge')",
             details = c("Both `tolerance$edge_edge` and `tolerance$line_line` were provided. Only `tolerance$edge_edge` will be considered.")
           )
          tolerance$line_line <- NULL
         }
```

# `which_longlat` argument

```{r}
if(!(which_longlat %in% c("sp", "sf"))){
        stop("The options for 'which_longlat' are 'sp' and 'sf'!")
      }
```

# `longlat` argument



```{r}
if(longlat){
        private$longlat <- TRUE
        private$which_longlat <- which_longlat
      }
```

# `proj4string` argument

```{r}
if(!is.null(proj4string)){
        if(!longlat){
          warning("proj4string was passed, so setting longlat to TRUE")
          longlat <- TRUE
          private$longlat <- TRUE
          private$which_longlat <- which_longlat
        }
        private$crs <- sf::st_crs(proj4string)
        private$proj4string <- proj4string
        crs <- private$crs
        private$transform <- !(sf::st_is_longlat(private$crs))
      }
```


# `crs` argument

```{r}
if(!is.null(crs)){
        if(!longlat){
          warning("crs was passed, so setting longlat to TRUE")
          longlat <- TRUE
          private$longlat <- TRUE
          private$which_longlat <- which_longlat          
        }        
        private$crs <- sf::st_crs(crs)
        private$proj4string <- sp::CRS(crs$input)
        proj4string <- private$proj4string
        private$transform <- !(sf::st_is_longlat(private$crs))        
      }

      if(longlat && (which_longlat == "sp") && is.null(proj4string)){
        proj4string <- sp::CRS("+proj=longlat +datum=WGS84")
        private$crs <- sf::st_crs(proj4string)
        private$proj4string <- proj4string
        private$transform <- !(sf::st_is_longlat(private$crs))        
      }

      if(longlat && (which_longlat == "sf") && is.null(crs)){
        crs <- sf::st_crs(4326)
        private$crs <- crs 
        private$transform <- !(sf::st_is_longlat(private$crs))        
      }

    # private$longlat <- longlat
```



# `vertex_unit` argument

```{r}
if((is.null(vertex_unit) && !is.null(length_unit)) || (is.null(length_unit) && !is.null(vertex_unit))){
      stop("If one of 'vertex_unit' or 'length_unit' is NULL, then the other must also be NULL.")
    }

    if(!is.null(vertex_unit)){
      vertex_unit <- vertex_unit[[1]]
      if(!is.character(vertex_unit)){
        stop("'vertex_unit' must be a string!")
      }
      if(!(vertex_unit %in% valid_units_vertex)){
        stop(paste("The possible options for 'vertex_unit' are ", toString(valid_units_vertex)))
      }
      private$vertex_unit <- vertex_unit
    }
```

# `lenght_unit` argument

```{r}
if(!is.null(length_unit)){
      length_unit <- length_unit[[1]]
      if(!is.character(length_unit)){
        stop("'length_unit' must be a string!")
      }
      if(length_unit == "degrees"){
        length_unit <- "km"
      }
      if(!(length_unit %in% valid_units_length)){
        stop(paste("The possible options for 'length_unit' are ", toString(valid_units_length)))
      }
      private$length_unit <- length_unit
    }

if(longlat){
      private$vertex_unit <- "degrees"
      if(!is.null(length_unit)){
        private$length_unit <- length_unit
      } else{
        private$length_unit <- "km"
      }
    } else if(!is.null(vertex_unit)){
        if(private$vertex_unit == "degrees"){
          longlat <- TRUE
          private$longlat <- TRUE
        }
    }

factor_unit <- process_factor_unit(private$vertex_unit, private$length_unit)
```


# `tolerance` argument again

```{r}
tolerance_default = list(vertex_vertex = 1e-7,
                             vertex_edge = 1e-7,
                              edge_edge = 0)



    for(i in 1:length(tolerance_default)){
      if(!(names(tolerance_default)[i] %in% names(tolerance))){
        tolerance[names(tolerance_default)[i]] <- tolerance_default[i]
      }
    }

    if(verbose > 0){
      message("Starting graph creation...")
      message(paste("LongLat is set to",longlat))
      if(longlat){
        message(paste("The unit for edge lengths is", private$length_unit))
        message(paste0("The current tolerances (in ",private$length_unit,") are:"))
        message(paste("\t Vertex-Vertex", tolerance$vertex_vertex))        
        message(paste("\t Vertex-Edge", tolerance$vertex_edge))           
        message(paste("\t Edge-Edge", tolerance$edge_edge))             
      } else{
        message("The current tolerances are:")
        message(paste("\t Vertex-Vertex", tolerance$vertex_vertex))        
        message(paste("\t Vertex-Edge", tolerance$vertex_edge))           
        message(paste("\t Edge-Edge", tolerance$edge_edge))    
      }
    }    

    if(is.null(tolerance$buffer_edge_edge)){
      tolerance$buffer_edge_edge <- max(tolerance$edge_edge/2 - 1e-10,0)
    }
    max_tol <- max(c(tolerance$vertex_vertex,
                     tolerance$vertex_edge,
                     tolerance$edge_edge))

    private$tolerance <- tolerance
```

# `V` and `E` arguments

---

`inherits()`

---

::: {.alert .alert-info}
Inside methods of the class, `self` refers to the object. Public members of the object are accessed with `self$x`, and assignment is done with `self$x <- y`. Whereas public members are accessed with `self`, like `self$x`, private members are accessed with `private`, like `private$x`.
:::


```{r}
PtE_tmp_edge_edge <- NULL
    PtE_tmp_edge_vertex <- NULL

    if(is.null(edges) && is.null(V) && is.null(E)) {
      edges <- logo_lines()
    }
    if(!is.null(edges)){
      if(!is.null(V) || !is.null(E)){
        warning("object initialized from edges, then E and V are ignored")
      }
      if (inherits(edges,"SpatialLinesDataFrame")) {
        tmp_lines = SpatialLines(edges@lines)
        self$edges <- lapply(1:length(tmp_lines), function(i){tmp_lines@lines[[i]]@Lines[[1]]@coords})
      } else if (inherits(edges,"SpatialLines")) {
        self$edges = lapply(1:length(edges), function(i){edges@lines[[i]]@Lines[[1]]@coords})
      } else if(inherits(edges, "MULTILINESTRING")) {
        coords_multilinestring <- sf::st_coordinates(edges)
        lines_ids <- unique(coords_multilinestring[,"L1"])
        self$edges <- lapply(1:length(lines_ids), function(i){coords_multilinestring[coords_multilinestring[,"L1"]==i ,1:2]})
      } else if(is.list(edges)){
        self$edges <- check_lines_input(edges)
      } else {
        stop("edges should either be a list, or of class MULTILINESTRING, SpatialLines or SpatialLinesDataFrame")
      }

    } else {
      if(is.null(V) || is.null(E)){
        stop("You must supply edges or V and E")
      }
      if(ncol(V)!=2 || ncol(E)!=2){
        stop("V and E must have two columns!")
      }
      edges <- list()
      for(i in 1:dim(E)[1]) {
        edges[[i]] <- rbind(V[E[i,1], ], V[E[i,2], ])
      }
      self$edges <- edges
    }
```


```{r, eval = TRUE}
x <- 5
class(x)
inherits(x, "numeric")  # Returns TRUE, because x is of class "numeric" or inherits from "numeric"
inherits(x, "integer")  # Returns FALSE, because x is not of class "integer"
is(x, "numeric")
is(x, "integer")
is.numeric(x)
is.integer(x)
```
- `inherits()` function is used to check whether an object inherits from a particular class or contains the attributes of a certain class. It returns a logical value (`TRUE` or `FALSE`) indicating whether the object inherits from the specified class. It behaves more or less like `is()` function.

# `set_first_weights()` function

```{r}
self$nE <- length(self$edges)
private$set_first_weights(weights = edge_weights)
```

- If everything is ok, `set_first_weights()` will internally do `private$edge_weights <- weights`. Press `Show` below to see the function's body.

```{r, class.source= "fold-hide"}
set_first_weights = function(weights = rep(1, self$nE)){
      if(is.null(weights)){
        weights <- rep(1, self$nE)
      }
    if(!is.vector(weights) && !is.data.frame(weights)){
      stop("'weights' must be either a vector or a data.frame!")
    }

    if(is.vector(weights)){
      if ( (length(weights) != 1) && (length(weights) != self$nE)){
        stop(paste0("The length of 'weights' must be either 1 or ", self$nE))
      }
      if(length(weights)==1){
        private$edge_weights <- rep(weights, self$nE)
      } else{
        private$edge_weights <- weights
      }
    } else{
      if(nrow(weights) != self$nE){
        stop("The number of rows of weights must be equal to the number of edges!")
      }
      private$edge_weights <- weights
    }

  }
```




```{r}
if(verbose > 0){
      message("Setup edges and merge close vertices")
    }

    t <- system.time(
      private$line_to_vertex(tolerance = tolerance$vertex_vertex,
                           longlat = private$longlat, factor_unit, verbose=verbose,
                           private$crs, private$proj4string, which_longlat, private$length_unit, private$vertex_unit,
                           project, which_projection, project_data)
      )

    if(verbose == 2){
      message(sprintf("time: %.3f s", t[["elapsed"]]))
    }


    if(length(self$edges) > 1){

    if (tolerance$edge_edge > 0) {
    private$addinfo <- TRUE

    if(verbose > 0){
      message("Find edge-edge intersections")
    }

    t <- system.time(
      points_add <- private$find_edge_edge_points(tol = tolerance$edge_edge, verbose=verbose,
      crs=private$crs, proj4string = private$proj4string, longlat=private$longlat, fact = factor_unit, which_longlat = which_longlat)
      )

    if(verbose == 2){
      message(sprintf("time: %.3f s", t[["elapsed"]]))
    }

    PtE <- points_add$PtE

    PtE[,2] <- PtE[,2]/self$edge_lengths[PtE[,1]]

    filter_tol <- ((PtE[,2] > max_tol/self$edge_lengths[PtE[,1]]) &
                     (PtE[,2] < 1- max_tol/self$edge_lengths[PtE[,1]]))

    PtE <- PtE[filter_tol,]

    if(!is.null(PtE)){
      if(nrow(PtE) == 0){
        PtE <- NULL
      }
    }

    if(!is.null(PtE)){
      if(verbose == 2){
        message(sprintf("Add %d new vertices", nrow(PtE)))
      }

      PtE <- na.omit(PtE)

      t <- system.time(
      private$add_vertices(PtE, tolerance = tolerance$edge_edge, verbose = verbose)
      )

      if(verbose == 2){
        message(sprintf("time: %.3f s", t[["elapsed"]]))
      }
    }

    private$clear_initial_info()
    }

    if(tolerance$vertex_edge > 0){
      private$addinfo <- TRUE
      if(verbose > 0){
        message("Snap vertices to close edges")
      }

      t <- system.time(
        PtE_tmp <- private$coordinates_multiple_snaps(XY = self$V,
                                              tolerance = tolerance$vertex_edge, verbose = verbose,
      crs=private$crs, proj4string = private$proj4string, longlat=private$longlat, fact = factor_unit, which_longlat = which_longlat)
        )

      if(verbose == 2){
        message(sprintf("time: %.3f s", t[["elapsed"]]))
      }
      edge_length_filter <- self$edge_lengths[PtE_tmp[,1]]

      filter_tol <- ((PtE_tmp[,2] > max_tol/edge_length_filter) &
                       (PtE_tmp[,2] < 1- max_tol/edge_length_filter))

      PtE_tmp <- PtE_tmp[filter_tol,,drop = FALSE]
      PtE_tmp <- unique(PtE_tmp)
      PtE_tmp <- PtE_tmp[order(PtE_tmp[,1], PtE_tmp[,2]),,drop = FALSE]

      if(!is.null(PtE_tmp)){
        if(nrow(PtE_tmp) == 0){
          PtE_tmp <- NULL
        }
      }

      if(!is.null(PtE_tmp)){
        if(verbose == 2){
          message(sprintf("Add %d new vertices", nrow(PtE_tmp)))
        }

        PtE_tmp <- na.omit(PtE_tmp)

        t <- system.time(
          private$add_vertices(PtE_tmp, tolerance = tolerance$vertex_edge, verbose=verbose)
          )

        if(verbose == 2){
          message(sprintf("time: %.3f s", t[["elapsed"]]))
        }
      }
      private$clear_initial_info()
    }

    if(merge_close_vertices){
      private$merge_close_vertices(factor_merge_close_vertices * tolerance$vertex_vertex, factor_unit)
    }

    if(is.logical(remove_circles)){
      if(remove_circles){
        private$remove_circles(tolerance$vertex_vertex, verbose=verbose,longlat = private$longlat, unit=length_unit, crs=private$crs, proj4string=private$proj4string, which_longlat=which_longlat, vertex_unit=vertex_unit, project_data)
      }
    } else {
        private$remove_circles(remove_circles, verbose=verbose,longlat = private$longlat, unit=length_unit, crs=private$crs, proj4string=private$proj4string, which_longlat=which_longlat, vertex_unit=vertex_unit, project_data)
        remove_circles <- TRUE
    }

    if(merge_close_vertices || remove_circles){
      if(verbose == 2){
        message("Recomputing edge lengths")
      }
      t <- system.time({
        self$edge_lengths <- private$compute_lengths(private$longlat, private$length_unit, private$crs, private$proj4string, private$which_longlat, private$vertex_unit, project_data,private$transform)
      })
       if(verbose == 2){
      message(sprintf("time: %.3f s", t[["elapsed"]]))
       }
    }
    # End of cond of having more than 1 edge
    }

    # Cleaning the edges

    if(verbose == 2){
      message("Post-processing the edges")
    }


    t <- system.time(
          self$edges <- lapply(self$edges, function(edge){
            tmp_edge <- edge[1:(nrow(edge)-1),]
            tmp_edge <- unique(tmp_edge)
            tmp_edge <- rbind(tmp_edge, edge[nrow(edge),,drop=FALSE])
            if(nrow(tmp_edge)>2){
              tmp_edge <- tmp_edge[2:nrow(tmp_edge),]
              tmp_edge <- unique(tmp_edge)
              tmp_edge <- rbind(edge[1,,drop=FALSE], tmp_edge)
            }
            rownames(tmp_edge) <- NULL
            return(tmp_edge)
          }
            )
    )

    if(verbose == 2){
          message(sprintf("time: %.3f s", t[["elapsed"]]))
    }

    # Checking if there is some edge with infinite length
    if(any(!is.finite(self$edge_lengths))){
      warning("There is at least one edge of infinite length. Please, consider redefining the graph.")
    }

    # Checking if there is some edge with zero length
    if(any(self$edge_lengths == 0)){
      warning("There is at least one edge of length zero. Please, consider redefining the graph.")
    }

    end_construction_time <- Sys.time()
    construction_time <- end_construction_time - start_construction_time

    if(verbose > 0){
      message(sprintf('Total construction time: %.2f %s', construction_time, units(construction_time)))

    }

    # Checking if graph is connected
    if (check_connected) {
      g <- graph(edges = c(t(self$E)), directed = FALSE)
      # components <- igraph::clusters(g, mode="weak")
      components <- igraph::components(g, mode="weak")
      nc <- components$no
      if(nc>1){
        message("The graph is disconnected. You can use the function 'graph_components' to obtain the different connected components.")
        private$connected = FALSE
      }
    }
    private$create_update_vertices()

    self$set_edge_weights(weights = private$edge_weights, kirchhoff_weights = private$kirchhoff_weights)


    if (remove_deg2) {
      if (verbose > 0) {
        message("Remove degree 2 vertices")
      }
      t <- system.time(
        self$prune_vertices(verbose = verbose)
      )
      if(verbose == 2){
        message(sprintf("time: %.3f s", t[["elapsed"]]))
      }
    }
    # Adding IDs to edges and setting up their class

    # for(i in 1:length(self$edges)){
    #   attr(self$edges[[i]], "id") <- i
    #   class(self$edges[[i]]) <- "metric_graph_edge"
    # }

    # Cloning the initial graph

    private$initial_graph <- self$clone()

    # Cloning again to add the initial graph to the initial graph
    private$initial_graph <- self$clone()

    # end of function initialize()
```



# `compute_geodist_PtE()` method

```{r}
compute_geodist_PtE = function(PtE,
                                 normalized = TRUE,
                                 include_vertices = TRUE, verbose = 0){
      if(verbose == 2){
        message("Processing the graph locations...")
      }
      t <- system.time({      
        graph.temp <- self$clone()
        graph.temp$clear_observations()
        df_temp <- data.frame(y = rep(0, dim(PtE)[1]),
                            edge_number = PtE[,1],
                            distance_on_edge = PtE[,2])
        if(sum(duplicated(df_temp))>0){
          warning("Duplicated locations were found when computing geodist. The returned values are given for unique locations.")
          df_temp <- unique(df_temp)
        }

        graph.temp$build_mesh(h = 10000)

        df_temp2 <- data.frame(y = 0, edge_number = graph.temp$mesh$VtE[1:nrow(self$V),1],
                                  distance_on_edge = graph.temp$mesh$VtE[1:nrow(self$V),2])

        df_temp$included <- TRUE
        temp_merge <- merge(df_temp, df_temp2, all = TRUE)

        df_temp$included <- NULL

        df_temp2 <- temp_merge[is.na(temp_merge["included"]),]

        nV_new <- sum(is.na(temp_merge["included"]))

        df_temp2$included <- NULL

        df_temp <- rbind(df_temp2, df_temp)

        df_temp[["__dummy"]] <- 1:nrow(df_temp)

        graph.temp$add_observations(data = df_temp,
                                     normalized = normalized,
                                     verbose=0,
                  suppress_warnings = TRUE)
      })

      if(verbose == 2){
        message(sprintf("time: %.3f s", t[["elapsed"]]))
      }

      if(verbose == 2){
        message("Turning observations of the auxiliary graph to vertices...")
      }      
      t <- system.time(
        graph.temp$observation_to_vertex(mesh_warning = FALSE)
      )
      if(verbose == 2){
        message(sprintf("time: %.3f s", t[["elapsed"]]))
       }
      if(verbose == 2){
        message("Creating auxiliary graph...")
      }
      t <- system.time({
        g <- graph(edges = c(t(graph.temp$E)), directed = FALSE)
        E(g)$weight <- graph.temp$edge_lengths
      })
      if(verbose == 2){
        message(sprintf("time: %.3f s", t[["elapsed"]]))
       }
      if(verbose>0){
        message("Computing geodesic distances...")
      }      
      t <- system.time(
        geodist_temp <- distances(g)
      )
      if(verbose == 2){
        message(sprintf("time: %.3f s", t[["elapsed"]]))
       }      

      if(length(graph.temp$PtV)[1]!=nrow(geodist_temp)){
        un_PtV <- unique(graph.temp$PtV)
        un_coords <- !duplicated(graph.temp$PtV)

        geodist_temp <- geodist_temp[un_PtV, un_PtV]

        tmp_vec <- graph.temp$.__enclos_env__$private$data[["__dummy"]][un_coords]

        un_ord <- order(tmp_vec)

        tmp_vec[un_ord] <- 1:length(tmp_vec)
        #Ordering back in the input order
        geodist_temp[tmp_vec,tmp_vec] <- geodist_temp
      } else{
          geodist_temp <- geodist_temp[graph.temp$PtV, graph.temp$PtV]
          #Ordering back in the input order
          geodist_temp[graph.temp$.__enclos_env__$private$data[["__dummy"]],graph.temp$.__enclos_env__$private$data[["__dummy"]]] <- geodist_temp
      }

      if(!include_vertices){
        geodist_temp <- geodist_temp[(nV_new+1):nrow(geodist_temp), (nV_new+1):nrow(geodist_temp)]
      }

      return(geodist_temp)
  }
```

# My change


```{r, eval = TRUE}
library(dplyr)
library(here)

# loading the data
load(here("Data_files/data_on_graph_with_covariates_no_consecutive_zeros_partialtomtom.RData"))
load(here("Graph_objects/graph_construction_28_03_2024partialtomtomwhichlonglatsf.RData"))
normalized = TRUE
aux = data_on_graph_with_covariates |>
  rename(distance_on_edge = .distance_on_edge, edge_number = .edge_number) |>
  as.data.frame() |>
  dplyr::select(edge_number, distance_on_edge, .group)

distmatrixlist = list()
PtE = aux %>% filter(.group == as.character(2)) %>% dplyr::select(-.group)


#distmatrixlist[[1]] = sf_graph$compute_geodist_PtE(PtE = PtEE, normalized = TRUE, include_vertices = FALSE)

self = sf_graph
graph.temp <- self$clone()
graph.temp$clear_observations()
df_temp <- data.frame(y = rep(0, dim(PtE)[1]),
                            edge_number = PtE[,1],
                            distance_on_edge = PtE[,2])
if(sum(duplicated(df_temp))>0){
  warning("Duplicated locations were found when computing geodist. The returned values are given for unique locations.")
df_temp <- unique(df_temp)
}
graph.temp$build_mesh(h = 10000)
aa = graph.temp$mesh$VtE
nvertices = sf_graph$nV
#df_temp2 contains all the vertices in graph coordinates
df_temp2 <- data.frame(y = 0, edge_number = graph.temp$mesh$VtE[1:nrow(self$V),1],
                                  distance_on_edge = graph.temp$mesh$VtE[1:nrow(self$V),2])
# creates a new column called included with entries all TRUE
df_temp$included <- TRUE
# just merge. entries in column included from df_temp is TRUE while from df_temp2 is NA
temp_merge <- merge(df_temp, df_temp2, all = TRUE)
# removes column included from df_temp
df_temp$included <- NULL
# this essentially does not change anything
df_temp2 <- temp_merge[is.na(temp_merge["included"]),]
nV_new <- sum(is.na(temp_merge["included"]))
df_temp2$included <- NULL
#df_temp is the same as temp_merge
df_temp <- rbind(df_temp2, df_temp)
df_temp[["__dummy"]] <- 1:nrow(df_temp)
jk = df_temp[,c(2,3)]
nrow(unique(jk))

graph.temp$add_observations(data = df_temp,
                                     normalized = normalized,
                                     verbose=0,
                  suppress_warnings = TRUE)
graph.temp$get_data() |> nrow()
graph.temp$observation_to_vertex(mesh_warning = FALSE)
graph.temp$nV
```

# Second attemp

```{r, eval = TRUE}
library(dplyr)
library(here)

# loading the data
load(here("Data_files/data_on_graph_with_covariates_no_consecutive_zeros_partialtomtom.RData"))
load(here("Graph_objects/graph_construction_28_03_2024partialtomtomwhichlonglatsf.RData"))
normalized = TRUE
# aux = data_on_graph_with_covariates |>
#   rename(distance_on_edge = .distance_on_edge, edge_number = .edge_number) |>
#   as.data.frame() |>
#   dplyr::select(edge_number, distance_on_edge, .group)
# 
# distmatrixlist = list()
# PtE = aux %>% filter(.group == as.character(2)) %>% dplyr::select(-.group)
#distmatrixlist[[1]] = sf_graph$compute_geodist_PtE(PtE = PtEE, normalized = TRUE, include_vertices = FALSE)

self = sf_graph
graph.temp <- self$clone()
graph.temp$clear_observations()
df_temp <- data.frame(y = rep(0, dim(PtE)[1]),
                            edge_number = PtE[,1],
                            distance_on_edge = PtE[,2])
if(sum(duplicated(df_temp))>0){
  warning("Duplicated locations were found when computing geodist. The returned values are given for unique locations.")
df_temp <- unique(df_temp)
}
graph.temp$build_mesh(h = 10000)
aa = graph.temp$mesh$VtE
nvertices = sf_graph$nV
#df_temp2 contains all the vertices in graph coordinates
df_temp2 <- data.frame(y = 0, edge_number = graph.temp$mesh$VtE[1:nrow(self$V),1],
                                  distance_on_edge = graph.temp$mesh$VtE[1:nrow(self$V),2])
# creates a new column called included with entries all TRUE
df_temp$included <- TRUE
# just merge. entries in column included from df_temp is TRUE while from df_temp2 is NA
temp_merge <- merge(df_temp, df_temp2, all = TRUE)
# removes column included from df_temp
df_temp$included <- NULL
# this essentially does not change anything
df_temp2 <- temp_merge[is.na(temp_merge["included"]),]
nV_new <- sum(is.na(temp_merge["included"]))
df_temp2$included <- NULL
#df_temp is the same as temp_merge
df_temp <- rbind(df_temp2, df_temp)
df_temp[["__dummy"]] <- 1:nrow(df_temp)
jk = df_temp[,c(2,3)]
nrow(unique(jk))

graph.temp$add_observations(data = df_temp,
                                     normalized = normalized,
                                     verbose=0,
                  suppress_warnings = TRUE)
graph.temp$get_data() |> nrow()
# graph.temp$observation_to_vertex(mesh_warning = FALSE)
# graph.temp$nV
private = list()
private$temp_PtE <- self$get_PtE()
n_group <- length(unique(private$data[[".group"]]))
l <- length(private$temp_PtE[, 1])
self$PtV <- rep(NA, l)
self$nE <- nrow(self$E)

```